# **Genomic Annotation Resources**

| Object Type                                              | Example Package Name                                           | Contents                                                                                                                                                                     |
|:---------------------------------------------------------|:---------------------------------------------------------------|:-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| OrgDb                                                    | [org.Hs.eg.db](http://bioconductor.org/packages/org.Hs.eg.db/) | Gene-based information for Homo sapiens; useful for mapping between gene IDs, Names, Symbols, GO and KEGG identifiers, etc.                                                  |
| [AnnotationHub](https://annotationhub.bioconductor.org/) | [AnnotationHub](https://annotationhub.bioconductor.org/)       | Provides a convenient interface to annotations from many different sources; objects are returned as fully parsed Bioconductor data objects or as the name of a file on disk. |

```{r eval=FALSE}
suppressMessages(library(AnnotationHub))
ah <- refreshHub(hubClass=c("AnnotationHub", "ExperimentHub"))
ah
save(ah,file = "data/AnnotationHub.Rdata")
```

```{r}
load(file = "data/AnnotationHub.RData")
ah
snapshotDate(ah)

# C:\Users\DELL\AppData\Local\R\cache\R\AnnotationHub
hubCache(ah)

# https://annotationhub.bioconductor.org
hubUrl(ah)
isLocalHub(ah)
dbconn(ah)
```

```{r eval=FALSE}
AnnotationHub::cache(ah[ah$dataprovider=="UCSC"|ah$dataprovider=="Ensembl"])
```

```{r}
length(ah)
ah$dataprovider |> unique()
```

```{r}
head(unique(ah$species))
head(unique(ah$rdataclass))
```

```{r}
query(ah, c("OrgDb"))
query(ah, c("^UCSC"))
```

```{r}
ah[ah$dataprovider=="UCSC"]
```
